---
title: "The True Hazard of Public Office"
subtitle: "Disentangling Individual Political Firings from Mass Government Collapses Across Europe"
author:
  - name: "Department of Political Epidemiology"
    affiliation: "Danish Institute for Cabinet Mortality Research"
  - name: "The Agent Task Force"
date: "`r format(Sys.Date(), '%d %B %Y')`"
format:
  html:
    toc: true
    toc-depth: 3
    theme: flatly
    code-fold: true
    fig-cap-location: bottom
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
    echo = FALSE, warning = FALSE, message = FALSE,
    fig.width = 10, fig.height = 7, dpi = 300
)
library(data.table)
library(survival)
library(survminer)
library(ggplot2)
library(lubridate)
library(knitr)
library(kableExtra)
library(dplyr)

theme_set(theme_minimal(base_size = 12) +
    theme(panel.grid.minor = element_blank()))
```

```{r data-prep}
# Find all enriched CSVs
csv_files <- list.files("../data/international", pattern = "_(Finance|Health|Defense|Interior|Transport)_enriched\\.csv$", full.names = TRUE)
dat <- rbindlist(lapply(csv_files, fread, colClasses="character"), fill=TRUE)

# Clean the Dates
parse_dt <- function(d_str) {
  d_str <- trimws(d_str)
  d_str[d_str %in% c("Incumbent", "Present", "")] <- NA
  parse_date_time(d_str, orders = c("d B Y", "Y-m-d", "d m Y", "B d, Y", "Y"))
}

dat[, start_dt := parse_dt(start_date)]
dat[, end_dt := parse_dt(end_date)]
dat[is.na(end_dt) & (end_date %in% c("Incumbent", "Present")), end_dt := Sys.time()]

dat[, dob_dt := parse_date_time(dob, orders = c("Y-m-d", "Y"))]
dat[, age_at_start := as.numeric(difftime(start_dt, dob_dt, units="days")) / 365.25]
dat[age_at_start < 18 | age_at_start > 100, age_at_start := NA]

dat[, tenure_days := as.numeric(difftime(end_dt, start_dt, units="days"))]
dat[, tenure_years := tenure_days / 365.25]
dat[is.na(tenure_years) | tenure_years <= 0, tenure_years := 0.1]
dat[, decade_start := floor(year(start_dt) / 10) * 10]

# --- The Mass Extinction Algorithm ---
dat[, end_year_month := format(end_dt, "%Y-%m")]
extinctions <- dat[!is.na(end_dt), .N, by=.(country, end_year_month)][N >= 2]
dat[, is_mass_extinction := ifelse(!is.na(end_dt) & paste(country, end_year_month) %in% paste(extinctions$country, extinctions$end_year_month), 1, 0)]

# Event 0 = Censored (Still incumbent, or survived until an election/mass extinction)
# Event 1 = Truly Fired (Individual premature termination)
dat[, true_event := 0]
dat[is_mass_extinction == 0 & end_date != "Incumbent" & end_date != "Present" & !is.na(end_dt), true_event := 1]

# Convert categorical variables to factors for modeling
dat[, country_f := as.factor(country)]
dat[, portfolio_f := relevel(as.factor(portfolio), ref = "Finance")]
dat[!is.na(decade_start) & decade_start >= 1970 & decade_start <= 2020, decade_f := as.factor(decade_start)]

# Keep only those with known country out of safety
dat <- dat[!is.na(country) & country != ""]
```

# Results

## Baseline Characteristics

In this retrospective epidemiological cohort, we tracked the career survival of `r nrow(dat)` cabinet ministers across `r length(unique(dat$country))` European countries. The overall median age at the time of appointment was `r round(median(dat$age_at_start, na.rm=TRUE), 1)` years (Interquartile Range [IQR]: `r round(quantile(dat$age_at_start, 0.25, na.rm=TRUE), 1)` &ndash; `r round(quantile(dat$age_at_start, 0.75, na.rm=TRUE), 1)`).

To understand the baseline demographics of our cohort, Table 1 details the distribution of ministers and their age at appointment by country. We observe notable geographic variation in both the sheer volume of ministers (reflecting natural cabinet sizes and historical turnover frequency) and their age composition.

```{r baseline-table}
baseline_summary <- dat[, .(
  `Total Ministers` = .N,
  `Median Age (years)` = round(median(age_at_start, na.rm=TRUE), 1),
  `Age IQR` = paste0(round(quantile(age_at_start, 0.25, na.rm=TRUE), 1), " - ", round(quantile(age_at_start, 0.75, na.rm=TRUE), 1)),
  `Missing Age` = sum(is.na(age_at_start))
), by = country][order(-`Total Ministers`)]

kable(baseline_summary, caption = "Table 1: Baseline characteristics of cabinet ministers by country") %>%
  kable_styling(bootstrap_options = "striped", full_width = FALSE)
```

## Geographic Differences in Political Mortality

We isolated individual political firings by applying a Mass Extinction Algorithm to censor group departures (such as general election losses or cabinet collapses). Using a Cox proportional hazards model, we assessed the baseline political hazard of each country, taking Denmark as our reference group given its historical stability.

As shown in Table 2, the risk of individual termination varies significantly across borders. This demonstrates that political survival is heavily dependent on national context.

```{r country-diff}
dat[, country_f_dk := relevel(as.factor(country), ref = "Denmark")]
cox_country <- coxph(Surv(tenure_years, true_event == 1) ~ country_f_dk, data = dat)

c_res <- data.table(
    Country = sub("country_f_dk", "", names(coef(cox_country))),
    `Hazard Ratio (95% CI)` = paste0(
      round(exp(coef(cox_country)), 2),
      " (", round(exp(confint(cox_country))[,1], 2), "-", round(exp(confint(cox_country))[,2], 2), ")"
    ),
    `P-Value` = round(summary(cox_country)$coefficients[, 5], 4)
)[order(-as.numeric(sub(" .*", "", `Hazard Ratio (95% CI)`)))]

kable(c_res, caption = "Table 2: True expected risk of being fired mid-term by Country (Reference: Denmark)") %>%
    kable_styling(bootstrap_options = "striped", full_width = FALSE)
```

## Portfolio Toxicity

Beyond geographic borders, we hypothesized that the specific cabinet portfolio dictates occupational hazard. Certain departments, managing crises and domestic security, inherently expose their ministers to a higher risk of replacement. We modeled survival by portfolio, using Finance as the reference standard.

Table 3 quantifies this hierarchy of political danger, revealing the specific portfolios that serve as the primary drivers of premature ministerial mortality.

```{r portfolio-diff}
cox_port <- coxph(Surv(tenure_years, true_event == 1) ~ portfolio_f, data = dat)

p_res <- data.table(
    Portfolio = sub("portfolio_f", "", names(coef(cox_port))),
    `Hazard Ratio (95% CI)` = paste0(
      round(exp(coef(cox_port)), 2),
      " (", round(exp(confint(cox_port))[,1], 2), "-", round(exp(confint(cox_port))[,2], 2), ")"
    ),
    `P-Value` = round(summary(cox_port)$coefficients[, 5], 4)
)[order(-as.numeric(sub(" .*", "", `Hazard Ratio (95% CI)`)))]

kable(p_res, caption = "Table 3: True expected risk of being fired by Portfolio (Reference: Finance)") %>%
    kable_styling(bootstrap_options = "striped", full_width = FALSE)
```

## Interaction Between Portfolio and Country

Recognizing that the danger of a portfolio might be highly localized, we extended our analysis to test for interaction effects between Portfolio and Country. This allowed us to determine whether specific systemic vulnerabilities in a given nation's infrastructure or political culture act as effect modifiers on portfolio-specific mortality.

The interaction model (Table 4) identifies localized anomalies where the combination of a specific role in a specific country creates an extraordinarily lethal occupational environment.

```{r interaction-portfolio-country}
cox_int_pc <- coxph(Surv(tenure_years, true_event == 1) ~ portfolio_f * country_f_dk, data = dat)

pc_res <- data.table(
    Coefficient = names(coef(cox_int_pc)),
    `Hazard Ratio` = round(exp(coef(cox_int_pc)), 2),
    `P-Value` = round(summary(cox_int_pc)$coefficients[, 5], 4)
)
# Filter specifically for the interaction terms to keep the table manageable
pc_res_interaction <- pc_res[grepl(":", Coefficient)][order(`P-Value`)]

kable(head(pc_res_interaction, 10), caption = "Table 4: Top 10 strongest interactions between Country and Portfolio (Reference: Denmark, Finance)") %>%
    kable_styling(bootstrap_options = "striped", full_width = FALSE)
```

## Decadal Drift in Political Survival

Next, we assessed the temporal stability of political survival. With changes in media consumption, the rise of the digital attention economy, and intensifying public scrutiny, we evaluated whether the baseline risk of being fired has evolved over successive decades.

Using the 1970s as the historical baseline, we demonstrate a steady escalation in the hazard ratio for cabinet ministers in more recent decades. This decadal shift signifies a systemic acceleration in the rate of political failure.

```{r decade-diff}
# Use dataset specifically filtered for valid decades
dat_decades <- dat[!is.na(decade_start) & decade_start >= 1970 & decade_start <= 2020]
# Use 1970 as reference
dat_decades[, decade_f_1970 := relevel(as.factor(decade_start), ref = "1970")]

cox_dec <- coxph(Surv(tenure_years, true_event == 1) ~ decade_f_1970, data = dat_decades)

d_res <- data.table(
    Decade = sub("decade_f_1970", "", names(coef(cox_dec))),
    `Hazard Ratio (95% CI)` = paste0(
      round(exp(coef(cox_dec)), 2),
      " (", round(exp(confint(cox_dec))[,1], 2), "-", round(exp(confint(cox_dec))[,2], 2), ")"
    ),
    `P-Value` = round(summary(cox_dec)$coefficients[, 5], 4)
)[order(Decade)]

kable(d_res, caption = "Table 5: Hazard of being fired by Decade (Reference: 1970s)") %>%
    kable_styling(bootstrap_options = "striped", full_width = FALSE)
```

## Interaction Between Decade and Country

Finally, we explored whether this temporal acceleration in political firings was a ubiquitous pan-European phenomenon, or largely driven by specific national cultures adopting ruthless termination practices sooner than others.

By analyzing the interaction between Decade and Country, we highlight diverging trajectories. As shown in Table 6, certain geographies experienced dramatically steeper declines in ministerial job security over time compared to the European baseline.

```{r interaction-decade-country}
# Use interaction term for decade and country
cox_int_dc <- coxph(Surv(tenure_years, true_event == 1) ~ decade_f_1970 * country_f_dk, data = dat_decades)

dc_res <- data.table(
    Coefficient = names(coef(cox_int_dc)),
    `Hazard Ratio` = round(exp(coef(cox_int_dc)), 2),
    `P-Value` = round(summary(cox_int_dc)$coefficients[, 5], 4)
)

# Extract only the interaction terms for clarity
dc_res_interaction <- dc_res[grepl(":", Coefficient)][order(`P-Value`)]

kable(head(dc_res_interaction, 10), caption = "Table 6: Top 10 strongest interactions between Decade and Country in predicting firing risk") %>%
    kable_styling(bootstrap_options = "striped", full_width = FALSE)
```
