---
title: "POLITICS KILLS: an actuarial analysis of life insurance limitations involving tragically important cabinet servants"
subtitle: "Ministerial mortality and the case for hazard pay in Danish government"
author:
  - name: "Department of Occupational Epidemiology"
    affiliation: "Danish Institute for Cabinet Mortality Research"
date: "`r format(Sys.Date(), '%d %B %Y')`"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3
    theme: flatly
    code_folding: hide
    fig_caption: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE,
                      fig.width = 10, fig.height = 7, dpi = 300)
library(data.table)
library(lubridate)
library(survival)
library(survminer)
library(ggplot2)
library(knitr)
library(kableExtra)
library(broom)

theme_set(theme_minimal(base_size = 12) +
          theme(panel.grid.minor = element_blank()))
```

# Abstract

**Objective** To determine whether Danish cabinet ministers receive adequate compensation for occupational hazards, including literal mortality.

**Design** Retrospective cohort study with actuarial life table comparison and occupational hazard benchmarking.

**Setting** Danish government offices, 1945-2022, where eight ministers ended their service permanently.

**Participants** 787 minister-government observations from 349 unique individuals serving in 37 Danish governments. Eight participants were unavailable for follow-up interviews due to death.

**Main outcome measures** Death in office; mid-term dismissal ("career death"); comparison to occupational fatality rates in fishing, logging, and other hazardous industries.

**Results** Three Prime Ministers died in office between 1942 and 1960, yielding an annualized fatality rate of 16,667 per 100,000. This rate exceeds commercial fishing (145 per 100,000) by a factor of 115. The overall ministerial death rate (415 per 100,000) exceeds fishing by 2.9-fold. Mid-term dismissal ("career death") affected 22.4% of ministers, producing an annual career fatality rate of 7.5%, which is 51 times the mortality rate in commercial fishing. Based on hazard pay principles applied to the fishing industry, Danish ministers should receive an additional $1,030,000 annually. Current ministerial salary ($227,000) covers only 22% of this implied premium.

**Conclusions** Danish ministers are substantially undercompensated for occupational hazards. The position of Prime Minister between 1942 and 1960 was more dangerous than any civilian occupation on record. We recommend actuarial review of ministerial life insurance policies and consideration of hazard pay supplements.

---

# What is already known on this topic

- Politicians as a group live longer than the general population (SMR 0.45-0.82)
- Commercial fishing has the highest occupational fatality rate among civilian professions (145 per 100,000)
- Hazard pay compensates workers for elevated mortality risk

# What this study adds

- Three Danish Prime Ministers died in office within 18 years (1942-1960), a rate 28 times the actuarial expectation
- Danish ministerial fatality rates exceed commercial fishing by 2.9-fold
- Mid-term dismissal ("career death") occurs 51 times more frequently than death in commercial fishing
- Current ministerial compensation covers only 22% of the implied hazard premium

---

# Introduction

The principle of hazard pay holds that workers facing elevated mortality risk deserve additional compensation. Commercial fishermen, logging workers, and aircraft pilots receive wage premiums reflecting their occupational hazards.^1^ No similar framework exists for political office.

We questioned whether this omission reflects genuinely low risk or merely failure to quantify ministerial hazards. Danish government records provide a natural experiment: complete enumeration of cabinet appointments from 1945 to 2022, including precise dates of service termination and, in eight cases, death.

Previous research established that politicians as a group enjoy favorable mortality compared with the general population, with standardized mortality ratios ranging from 0.45 in Italy to 0.82 in New Zealand.^2^ However, a separate BMJ study found that elected heads of government lived 4.4 fewer years than losing candidates, with a hazard ratio of 1.23.^3^ This paradox suggests that while political involvement may be protective, leadership itself carries costs.

We hypothesized that Danish ministerial mortality rates, properly calculated, would exceed those of recognized hazardous occupations. We further hypothesized that mid-term dismissal ("career death") represents an additional occupational hazard warranting compensation.

---

# Methods

```{r load-data}
# Load Danish ministerial data
dk_data <- fread("../data/danish_cabinets.csv", encoding = "Latin-1")
dk_data[, V1 := NULL]
Encoding(dk_data$ministerpost) <- "latin1"

# Parse dates and calculate tenure
dk_data[, start := as.Date(start)]
dk_data[, stop := as.Date(stop)]
dk_data[, tenure_days := as.numeric(difftime(stop, start, units = "days"))]
dk_data[, tenure_years := tenure_days / 365.25]

# Identify deaths (marked with †)
dk_data[, died_in_office := grepl("†", ministerpost)]

# Classify ministry type
extract_ministry <- function(x) {
  x <- tryCatch(tolower(x), error = function(e) x)
  if (is.na(x) || x == "") return("Other")
  if (grepl("statsminister|premierminister", x)) return("Prime Minister")
  if (grepl("finansminister", x)) return("Finance")
  if (grepl("indenrigsminister", x)) return("Interior")
  if (grepl("landbrug", x)) return("Agriculture")
  return("Other")
}
dk_data[, ministry := sapply(ministerpost, extract_ministry)]

# Government-level calculations
dk_data[, govt_start := min(start, na.rm = TRUE), by = regering]
dk_data[, govt_end := max(stop, na.rm = TRUE), by = regering]
dk_data[, person_end := max(stop, na.rm = TRUE), by = .(regering, navn)]

# Create analysis dataset
analysis <- dk_data[year(start) >= 1945 & !is.na(tenure_days) & tenure_days > 0]

# Person-government aggregation
person_govt <- analysis[, .(
  start_date = min(start),
  end_date = max(stop),
  total_tenure = sum(tenure_years, na.rm = TRUE),
  ministry = ministry[which.min(start)],
  govt_end = govt_end[1]
), by = .(navn, regering)]

person_govt[, days_before_govt_end := as.numeric(difftime(govt_end, end_date, units = "days"))]
person_govt[, dismissed := as.numeric(days_before_govt_end > 30)]
```

## Study design

We conducted a retrospective cohort study of all ministerial appointments in Denmark from 7 November 1945 to 1 March 2022. We obtained records from official government sources, which mark deceased ministers with the dagger symbol (†).

## Outcome definitions

**Death in office**: Minister died while serving, identified by † notation.

**Career death**: Mid-term dismissal, defined as departure more than 30 days before government termination.

## Actuarial comparison

We compared observed deaths to expected deaths using Danish life tables. We assumed an average ministerial age of 55-60 years, corresponding to an annual mortality probability of 0.5%.^4^

## Occupational hazard benchmarking

We compared ministerial fatality rates to published occupational fatality rates from the US Bureau of Labor Statistics.^5^ We calculated fatality rates per 100,000 full-time equivalent workers per year.

## Hazard pay estimation

We estimated the implied hazard pay premium by applying the wage differential observed in commercial fishing (approximately $20,000 annually) proportionally to the relative risk.

---

# Results

## Deaths in office

```{r deaths-analysis}
# Find all deaths
deaths <- dk_data[died_in_office == TRUE]
n_deaths <- uniqueN(deaths$navn)

# Death details
death_details <- deaths[, .(
  Position = fifelse(any(grepl("statsminister", tolower(ministerpost))), "Prime Minister",
             fifelse(any(grepl("finansminister", tolower(ministerpost))), "Finance Minister",
             fifelse(any(grepl("indenrigsminister", tolower(ministerpost))), "Interior Minister",
             "Minister"))),
  Year = year(max(stop))
), by = .(Name = navn)][order(Year)]

# Prime Minister deaths
pm_deaths <- death_details[Position == "Prime Minister"]
```

Eight ministers died in office between 1942 and 2022 (Table 1). Three were Prime Ministers: Thorvald Stauning (1942), Hans Hedtoft (1955), and H.C. Hansen (1960).

```{r table1-deaths}
kable(death_details,
      col.names = c("Name", "Position", "Year"),
      caption = "Table 1. Danish ministers who died in office") %>%
  kable_styling(bootstrap_options = c("striped")) %>%
  column_spec(1, bold = TRUE)
```

The three Prime Minister deaths occurred within an 18-year interval, yielding an annualized rate of one death per 6 years. Hans Hedtoft died of a heart attack during a Nordic Council meeting in Stockholm on 29 January 1955.

## Actuarial analysis

```{r actuarial}
# Parameters
pm_deaths_n <- 3
pm_period_years <- 18  # 1942-1960
expected_annual_mortality <- 0.005  # 0.5% for 55-60 year-old men

# Expected deaths for 1 PM over 18 years
expected_pm_deaths <- pm_period_years * 1 * expected_annual_mortality
observed_pm_deaths <- pm_deaths_n
smr_pm <- observed_pm_deaths / expected_pm_deaths

# Overall ministers
years_total <- 77  # 1945-2022
ministers_at_risk <- 25  # approximate average
expected_minister_deaths <- years_total * ministers_at_risk * expected_annual_mortality
observed_minister_deaths <- n_deaths
smr_ministers <- observed_minister_deaths / expected_minister_deaths
```

For Prime Ministers between 1942 and 1960, the expected number of deaths based on Danish life tables was `r round(expected_pm_deaths, 2)`. We observed `r observed_pm_deaths` deaths, yielding a standardized mortality ratio of `r round(smr_pm, 1)` (Table 2).

For all ministers between 1945 and 2022, assuming 25 ministers at risk and average age 55, the expected number of deaths was `r round(expected_minister_deaths, 1)`. We observed `r observed_minister_deaths` deaths, yielding a standardized mortality ratio of `r round(smr_ministers, 2)`.

```{r table2-actuarial}
actuarial_table <- data.table(
  Group = c("Prime Ministers (1942-1960)", "All ministers (1945-2022)"),
  `Period (years)` = c(18, 77),
  `Expected deaths` = c(round(expected_pm_deaths, 2), round(expected_minister_deaths, 1)),
  `Observed deaths` = c(observed_pm_deaths, observed_minister_deaths),
  SMR = c(round(smr_pm, 1), round(smr_ministers, 2))
)

kable(actuarial_table,
      caption = "Table 2. Standardized mortality ratios for Danish ministers") %>%
  kable_styling(bootstrap_options = c("striped", "hover")) %>%
  row_spec(1, bold = TRUE, background = "#f8d7da")
```

## Occupational hazard comparison

```{r occupation-comparison}
# Calculate fatality rates per 100,000
pm_fatality_rate <- (pm_deaths_n / pm_period_years) * 100000  # per PM-year
minister_fatality_rate <- (observed_minister_deaths / years_total / ministers_at_risk) * 100000

occupation_data <- data.table(
  Occupation = c(
    "Danish Prime Minister (1942-1960)",
    "Danish cabinet minister (1945-2022)",
    "Commercial fishing",
    "Logging",
    "Aircraft pilots",
    "Roofers",
    "Garbage collectors",
    "Truck drivers",
    "Construction (all)",
    "All US workers"
  ),
  `Fatality rate` = c(
    round(pm_fatality_rate, 0),
    round(minister_fatality_rate, 0),
    145, 82, 58, 47, 34, 24, 10, 3.5
  )
)

occupation_data[, `Ratio to fishing` := round(`Fatality rate` / 145, 1)]
```

Danish Prime Ministers between 1942 and 1960 experienced a fatality rate of `r format(round(pm_fatality_rate, 0), big.mark=",")` per 100,000 person-years (Table 3). This rate exceeds commercial fishing by a factor of `r round(pm_fatality_rate / 145, 0)`.

```{r table3-occupations}
kable(occupation_data,
      col.names = c("Occupation", "Fatality rate (per 100,000)", "Ratio to fishing"),
      caption = "Table 3. Occupational fatality rates compared with Danish ministerial positions") %>%
  kable_styling(bootstrap_options = c("striped", "hover")) %>%
  row_spec(1:2, bold = TRUE, background = "#f8d7da") %>%
  row_spec(3, background = "#fff3cd")
```

```{r figure1-occupation, fig.cap="Figure 1. Occupational fatality rates per 100,000 workers. Danish Prime Ministers (1942-1960) shown truncated; actual rate is 16,667 per 100,000. Commercial fishing, conventionally the most dangerous civilian occupation, shown in yellow for reference."}
plot_data <- occupation_data[Occupation != "Danish Prime Minister (1942-1960)"]
plot_data[, Occupation := factor(Occupation, levels = Occupation[order(`Fatality rate`)])]

# Add color coding
plot_data[, Category := fifelse(grepl("Danish", Occupation), "Minister",
                         fifelse(grepl("fishing", Occupation), "Fishing", "Other"))]

ggplot(plot_data, aes(x = `Fatality rate`, y = Occupation, fill = Category)) +
  geom_col() +
  geom_text(aes(label = `Fatality rate`), hjust = -0.2, size = 3.5) +
  scale_fill_manual(values = c("Minister" = "#dc3545", "Fishing" = "#ffc107", "Other" = "#6c757d")) +
  labs(
    title = "Occupational fatality rates per 100,000 workers",
    subtitle = "Danish PM (1942-1960) rate of 16,667 not shown (would extend 115x beyond fishing)",
    x = "Annual fatalities per 100,000 FTE workers",
    y = NULL
  ) +
  theme(legend.position = "none") +
  xlim(0, 500)
```

## Career mortality

```{r career-mortality}
n_obs <- nrow(person_govt)
n_dismissed <- sum(person_govt$dismissed, na.rm = TRUE)
dismissal_rate <- n_dismissed / n_obs
annual_career_death_rate <- dismissal_rate / 3  # assuming 3-year average tenure

# Compare to fishing
fishing_mortality <- 145 / 100000
career_vs_fishing <- annual_career_death_rate / fishing_mortality
```

Of `r n_obs` minister-government observations, `r n_dismissed` (`r round(dismissal_rate * 100, 1)`%) ended in mid-term dismissal. Assuming average tenure of 3 years, this yields an annual "career death" rate of `r round(annual_career_death_rate * 100, 1)`%.

This career mortality rate is `r round(career_vs_fishing, 0)` times the literal mortality rate in commercial fishing (Figure 2).

```{r figure2-career, fig.cap="Figure 2. Annual 'mortality' rates by type. Career death (ministerial dismissal) shown alongside literal death rates. Note logarithmic scale."}
mortality_comparison <- data.table(
  Type = c("Career death (ministers)", "Literal death (fishing)",
           "Literal death (ministers)", "Literal death (all workers)"),
  Rate = c(annual_career_death_rate * 100, 0.145, minister_fatality_rate/1000, 0.0035),
  Category = c("Career", "Literal", "Literal", "Literal")
)

ggplot(mortality_comparison, aes(x = reorder(Type, Rate), y = Rate, fill = Category)) +
  geom_col() +
  geom_text(aes(label = sprintf("%.2f%%", Rate)), hjust = -0.1, size = 4) +
  coord_flip() +
  scale_y_log10(labels = function(x) paste0(x, "%")) +
  scale_fill_manual(values = c("Career" = "#dc3545", "Literal" = "#6c757d")) +
  labs(
    title = "Annual 'mortality' rates: career vs literal",
    x = NULL,
    y = "Annual rate (%, log scale)"
  ) +
  theme(legend.position = "none") +
  ylim(0.001, 15)
```

## Hazard pay calculation

```{r hazard-pay}
# Current salaries (2025, from regeringen.dk)
pm_salary_dkk <- 1958176
minister_salary_dkk <- 1566541
dkk_to_usd <- 1 / 7.46 * 1.08  # via EUR

pm_salary_usd <- pm_salary_dkk * dkk_to_usd
minister_salary_usd <- minister_salary_dkk * dkk_to_usd

# Fishing hazard premium estimate
fishing_hazard_premium <- 20000  # USD

# Implied premium based on career death rate
implied_premium <- fishing_hazard_premium * career_vs_fishing
coverage_ratio <- minister_salary_usd / implied_premium
```

Commercial fishermen receive an estimated hazard premium of approximately $20,000 annually.^6^ Applying this rate proportionally to the ministerial career death risk (`r round(career_vs_fishing, 0)` times fishing mortality) implies a premium of $`r format(round(implied_premium, 0), big.mark = ",")` annually.

Current ministerial salary is $`r format(round(minister_salary_usd, 0), big.mark = ",")`, representing `r round(coverage_ratio * 100, 0)`% of the implied hazard premium (Table 4).

```{r table4-compensation}
comp_table <- data.table(
  Measure = c(
    "Current ministerial salary",
    "Implied hazard premium (career death)",
    "Shortfall",
    "Coverage ratio"
  ),
  Value = c(
    paste0("$", format(round(minister_salary_usd, 0), big.mark = ",")),
    paste0("$", format(round(implied_premium, 0), big.mark = ",")),
    paste0("$", format(round(implied_premium - minister_salary_usd, 0), big.mark = ",")),
    paste0(round(coverage_ratio * 100, 0), "%")
  )
)

kable(comp_table, col.names = c("Measure", "Amount (USD)"),
      caption = "Table 4. Ministerial compensation versus implied hazard premium") %>%
  kable_styling(bootstrap_options = c("striped")) %>%
  row_spec(3, bold = TRUE, background = "#f8d7da")
```

---

# Discussion

## Principal findings

Danish Prime Ministers between 1942 and 1960 faced mortality risk 115 times that of commercial fishing, conventionally the most dangerous civilian occupation. The overall ministerial death rate exceeded fishing by 2.9-fold. When career termination is treated as an occupational hazard, ministerial risk exceeds fishing mortality by 51-fold.

Current ministerial compensation covers approximately 22% of the implied hazard premium. Danish ministers are thus substantially undercompensated relative to established principles of occupational risk pricing.

## The leadership mortality paradox

Our findings align with the broader literature on political mortality. Politicians as a group enjoy favorable survival compared with the general population, likely reflecting socioeconomic advantages, access to healthcare, and selection effects.^2^ However, those who achieve leadership positions face elevated mortality compared with unsuccessful candidates.^3^

This paradox suggests that leadership itself, rather than political involvement generally, carries health costs. The three Danish Prime Ministers who died in office between 1942 and 1960 support this interpretation.

## Why were the 1940s-1960s so deadly?

The concentrated mortality of Danish Prime Ministers during this period may reflect several factors:

**Wartime stress**: Thorvald Stauning governed during the German occupation and died in 1942. The strain of maintaining Danish institutions under occupation plausibly accelerated his decline.

**Post-war demands**: Hans Hedtoft and H.C. Hansen led Denmark through NATO membership, welfare state expansion, and Cold War tensions. Both died of cardiovascular causes consistent with chronic stress.

**Selection effects**: This generation of leaders entered politics young and served extended tenures, accumulating stress exposure.

## Limitations

Our actuarial comparison assumes ministers face population-average mortality absent their position. Ministers likely differ from the general population in education, income, and health behaviors, potentially biasing our SMR estimates.

The hazard pay calculation treats career termination as equivalent to occupational death. This analogy, while illustrative, may overstate the harm of dismissal. Former ministers frequently return to lucrative private sector positions.

Sample sizes for Prime Minister mortality are small (n=3 deaths), producing wide confidence intervals around point estimates.

## Policy implications

If society wishes to attract talent to ministerial positions, compensation should reflect occupational hazards. Our analysis suggests a hazard premium of approximately $800,000 annually would achieve parity with established risk pricing.

Alternatively, occupational health interventions could reduce ministerial mortality. Mandatory cardiovascular screening, stress management programs, and limits on working hours might improve survival. Hans Hedtoft's fatal heart attack at a Nordic Council meeting suggests that even international cooperation can be hazardous.

## Conclusions

Danish ministerial positions carry substantial occupational hazards that are not reflected in current compensation. The position of Prime Minister between 1942 and 1960 was more dangerous than any civilian occupation on record. We recommend actuarial review of ministerial life insurance policies and consideration of hazard pay supplements reflecting true occupational risk.

Ministers contemplating their career prospects should note that their annual probability of "death" (dismissal) exceeds that of commercial fishermen by a factor of 51. Those aspiring to lead should further note that three of their predecessors died in office within living memory.

---

# References

1. Bureau of Labor Statistics. Census of Fatal Occupational Injuries Summary, 2023. Washington, DC: US Department of Labor, 2024.

2. Clarke PM, Schluter PJ, Grayson M, et al. The comparative mortality of an elite group in the long run of history: an observational analysis of politicians from 11 countries. *Eur J Epidemiol* 2022;37:1-9.

3. Olenski AR, Abola MV, Jena AB. Do heads of government age more quickly? Observational study comparing mortality between elected leaders and runners-up in national elections of 17 countries. *BMJ* 2015;351:h6424.

4. Statistics Denmark. Life tables, 2023. Copenhagen: Statistics Denmark, 2024.

5. Bureau of Labor Statistics. Injuries, Illnesses, and Fatalities. Available at: https://www.bls.gov/iif/

6. Kniesner TJ, Viscusi WK. The value of a statistical life. *Vanderbilt Law Research Paper* 2019;19-15.

---

# Funding

None. This study received no specific funding. The authors conducted it between meetings of uncertain survival probability.

# Competing interests

None declared. The authors hold no ministerial appointments and, having reviewed the evidence, have no plans to seek one.

# Ethical approval

Not required. All subjects were public figures. Eight are deceased.

# Data sharing

Data and analysis code are available from the corresponding author. Life insurance applications available upon request.

---

*Submitted for consideration to the BMJ Christmas Issue 2024*

